Homeless Populations in Toronto

### 2022-02-05 outputs/paper/paper.Rmd
updates

```{r, include=FALSE}
# install.packages("opendatatoronto")
library(opendatatoronto)
library(tidyverse)
library(tidyr)
library(knitr)
library(janitor)
library(lubridate)
```

@@

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

@@

>>>>

```{r, include = FALSE}
#install.packages("opendatatoronto")
library(opendatatoronto)
library(tidyverse)
library(tidyr)
library(knitr)
library(janitor)
library(lubridate)
library(tibble)
library(dplyr)
library(scales)
```

@@

@@

## Data Characteristics ##

The first aspect of the data I investigated relates to the attrition rate of housing, ie; the likelihood that a homeless person moved to permanent housing does not return to the shelter. Figure \@ref(fig:housingattrition) graphs the number of people moved into permanent housing against the number of people who have returned to a shelter from after being moved to permanent housing. There has been a sharp decrease in the number of people being moved to housing since the beginning of 2020, while the number of people returning to shelters after being placed in permanent housing has remained relatively stable. This indicates a sharp decrease in the attrition rates of housing placements, which is one of the primary gauges of how successful Toronto's homeless services are at eliminating homelessness.

```{r, include=FALSE}
populations <- read.csv(here::here("inputs/data/raw_data.csv"))

populations$date.mmm.yy. <- Map(paste, populations$date.mmm.yy., "-01", sep="")
populations$date.mmm.yy. <- strptime(populations$date.mmm.yy., "%b-%y-%d")
populations$date.mmm.yy. <- as.Date(populations$date.mmm.yy., format = "%Y-%m-%d")
names(populations)[2] <- "Date"

homeless_housing <-
  populations %>%

  filter(population_group == 'All Population') %>%   # Filter data so that we only look at the full population for each month

  select(Date, returned_from_housing, moved_to_housing) %>%

  mutate(Attrition = percent(1 - returned_from_housing / moved_to_housing)) %>%

  rename("Returned from Housing" = returned_from_housing,   # Rename columns so that table reads nicer
         "Moved to Housing" = moved_to_housing)
```

```{r housingattrition, fig.cap = 'Moved to Housing vs Returned from Housing', fig.width = 8, fig.height = 4, echo = FALSE, message = FALSE, warning = FALSE, out.width = '80%', fig.align = "center"}

graph_housing <-
  homeless_housing %>% 
  tidyr::gather("Moved to Housing vs Returned from Housing", "Population", 2:3, na.rm = FALSE, convert = FALSE)

graph_housing %>% 
  ggplot(aes(x = Date, y = Population, color = `Moved to Housing vs Returned from Housing`, group = 2)) +
  geom_point() +
  geom_line(aes(group = `Moved to Housing vs Returned from Housing`))
```
